---
title: "R Notebook"
output: html_notebook
---
---
title: "Homework 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

###Import and Tidy
Read the gazetteer data as-is (all columns; no type conversion) into a gaz_raw tibble.

```{r uploaddata, message = FALSE}
library(readr)
library(tidyverse)


california_raw <- read_delim("CA_Features_20180401.txt", 
    "|", col_names = TRUE)













```

Copy only the following columns into a gaz tibble (you can rename them if you like):

```{r COPY}

california_select <- select(california_raw, FEATURE_ID, FEATURE_NAME, FEATURE_CLASS, STATE_ALPHA, COUNTY_NAME, PRIM_LAT_DEC, PRIM_LONG_DEC, SOURCE_LAT_DEC, SOURCE_LONG_DEC, ELEV_IN_M, MAP_NAME, DATE_CREATED, DATE_EDITED)

```

Convert the gaz columns to the appropriate type. Convert any placeholders for unknown data to NA

```{r selectingdata, warning = FALSE}

california_convert <- type_convert(california_select, col_types = cols(
        DATE_CREATED = col_date(format = "%m/%d/%Y"), 
        DATE_EDITED = col_date(format = "%m/%d/%Y"), 
        STATE_NUMERIC = col_integer()))





```

Delete from gaz rows where:
the primary latitude or longitude are unknown
the feature is not in California

```{r deleting}

cali_ds <- filter(california_convert, STATE_ALPHA == "CA")
cali_ds2 <- filter(cali_ds, PRIM_LAT_DEC != "NA", PRIM_LAT_DEC != 0, PRIM_LONG_DEC != "NA", PRIM_LONG_DEC != 0)

```

Write the gaz tibble to a CSV file (using "|" as a delimiter).
```{r csv}

write_delim(cali_ds2,  delim = "|", "calselections.csv")


```


##Analyze

What is the most-frequently-occuring feature name?
Church of Christ
```{r name}


cali_names <- cali_ds2 %>% 
  group_by(FEATURE_NAME) %>% 
  summarize(freq = n()) %>% 
  filter(freq == max(freq))


cali_names  

```

What is the least-frequently-occuring feature name?
Isthmus, Sea

```{r least}
cali_features <- cali_ds2 %>% 
  group_by(FEATURE_CLASS) %>% 
  summarize(freq = n()) %>% 
  filter(freq == min(freq))

cali_features
```

What is the approximate center point of each county?

```{r center}

county <- cali_ds2 %>% 
  select(COUNTY_NAME, PRIM_LAT_DEC, PRIM_LONG_DEC) %>% 
  group_by(COUNTY_NAME) %>% 
  summarize(averagelong = mean(PRIM_LONG_DEC), averagelat = mean(PRIM_LAT_DEC))
  

county
```

What are the fractions of the total number of features in each county that are natural? man-made?

```{r fractions, warning = FALSE, message=FALSE}
props <- cali_ds2 %>% 
  group_by(FEATURE_CLASS) %>% 
  select(FEATURE_CLASS) %>% 
  summarize(freq = n())
  
write.csv(props, "calprops.csv")
natural <- read_csv("calprops_edited.csv")
natural_select <- select(natural, FEATURE_CLASS, NATURAL, MANMADE, TOTALFEATURE)


california_props <- left_join(cali_ds2, natural_select, by = "FEATURE_CLASS")


california_props_edit <- california_props %>% 
  select(COUNTY_NAME, NATURAL, MANMADE, TOTALFEATURE) %>% 
  group_by(COUNTY_NAME)


california_props_edit2 <- california_props %>% 
  select(COUNTY_NAME, NATURAL, MANMADE, TOTALFEATURE) %>% 
  group_by(COUNTY_NAME) %>% 
  summarize(totnat=sum(NATURAL), totmm=sum(MANMADE), total=sum(TOTALFEATURE))

county_props <- california_props_edit2 %>% 
  mutate(propnat = round(totnat/total, digits = 3), propmm = round(totmm/total, digits = 3))

county_props_final <- county_props %>% 
  select(COUNTY_NAME, propnat, propmm)

county_props_final




 
```

